Genetic Dilution erodes productivity: Exploring farmers' low adoption levels of improved maize in Ethiopia

### 2023-03-23 paper.tex

% 	\affil[1]{\small \emph{The World Bank}}
% 	\affil[2]{\small \emph{FAO}}
% 	\affil[3]{\small \emph{Assistant Research Professor at the Center for Data Science for Enterprise and Society, Cornell University}}
\author{%
 Aleksandr Michuda \footnote{Center for Data Science for Enterprise and Society, Cornell University}%
 \and Cristina Chiarella\footnote{UCLouvain}%
 \and Oscar Barriga-Cabanillas\footnote{The World Bank}%
  \and Juan Sebastian Correa \footnote{FAO}%
  }
%   Assistant Research Professor at the 
	\date{\today}

    \clearpage\maketitle
	\thispagestyle{empty}
	\vspace*{-2em}
    \begin{center}\begin{abstract}
    \singlespacing
			\noindent  
			Despite large investments in  developing new  maize varieties, evidence shows that few farmers consistently adopt this new technology over time. We study the adoption trajectories of farmers over a three-round panel data set to explain low adoption rates with heterogeneous returns to adoption. By relying on self-reported use of improved seeds, we find a puzzling answer for why farmers do not use new seed varieties: adoption leads to seemingly small and negative returns. Heterogeneity analysis across different levels of comparative advantage show that farmers who adopt do not have a comparative advantage in adopting improved seed, nor do they exhibit high returns to adopting. To reconcile this result with the expected higher yields for improved seeds, we exploit unique information on maize DNA-fingerprinting collected over the same areas in 2018. We find that the negative returns to adoption mask heterogeneity in the actual quality of the genetic material of the seeds farmers categorized as improved varieties in the self-reported data. Positive returns to adoption are found for those farmers that use higher-purity germplasm, drought-tolerant maize, and newly released varieties. When we account for misclassification in our heterogeneity analysis, we find that the puzzle is resolved. Our findings point to the wide dispersion of older and genetically diluted varieties, for which poorer farmers may be paying a premium. The implications of our findings speak to the need for policies to better target context and geography, expand accessibility of improved seeds, and make higher yields varieties more inclusive.

@@

%%% Why it is important
Food security is a major challenge in many East African countries, and Ethiopia is no exception, having historically struggled to provide an adequate and reliable food supply \citep{Ramakrishna2002-hv, Jaleta2018-oj}. Since the drought of 1984, Ethiopia has become increasingly reliant on maize, and as of today, maize constitutes the most widespread crop in the country, with total area cultivated increasing and average yields almost doubling since 1990.\footnote{FAO statistics show the total area cultivated with Maize in 2019 is 25\% larger than the next most widespread crop (sorghum), and 27\% and 139\% larger than the next two following crops, wheat and barley, respectively.}\footnote{Improved security conditions and maize productivity contributed with an important decrease in undernourishment which fell from 47\% in 1990 to 14\% in 2018 (The World Bank, 2021). But despite these gains, food insecurity still affects more than half the population of Ethiopia. The country has also recently seen increases in food insecurity because of COVID-19, the conflict in Ethiopia’s Tigray region and the desert locust outbreak.} But despite this, progress a large productivity gap persists between the African continent and the rest of the developing world.\footnote{From 2016 to the present, cereal crop yields in Sub-Saharan Africa have consistently been around 70-80\% lower than Europe or North America \citep{worldbankyield}.} The persistence of this gap is often attributed to different adoption rates of green revolution technologies, which include improved seeds. Evidence has shown that the adoption of improved seed leads to increases in agricultural yields \citep{Carter2014-fm}, food security \citep{Shiferaw2014-op}, and poverty reduction \citep{Minten2008-tj}.\footnote{Nevertheless, critics of green revolution technologies have also claimed that these technological packages have not been as profitable for many farmers in Africa as promised. Even the most successful of these countries, Ethiopia, is not on track to meet the goals of the Alliance for a Green Revolution for Africa (AGRA). This is due to high costs, and heavy dependence on subsidies \citep{Wise20}.} But in fact, only 13\% of maize farmers in Ethiopia have adopted improved seeds consistently between 2011 and 2015. A large percentage has adopted new varieties inconsistently, and the largest share of farmers (62\%) have not used improved seeds at all during the same period.

@@

Several answers have been proposed to explain such low adoption of improved varieties. These include lack of liquidity and other economic constraints \citep{Carter2014-fm}, growing conditions such as the type of soil, input use, and farmer characteristics \citep{Munshi2004-og}, imperfections in credit markets \citep{Croppenstedt2003-pq}, ambiguously defined property rights \citep{Place2000-el}, lack of commitment devices \citep{Duflo2009-iv}, high transportation costs that increase the cost of agricultural inputs \citep{Byerlee2013-qk}, and improved seed supply constraints \citep{Bird2020-nt}. An alternative mechanism for a lack of improvement has the misclassification of seed and seed dilution. Net returns to improved seed is often lower than expected in the long term because a large percentage of farmers use diluted improved seed varieties, or ones that have been recycled over many generations. Evidence suggests that yield improvement and disease resistance wanes over time \citep{Abate2015-rj}. Conversely, given the ubiquity of improved seed varieties over the course of decades of access in Ethiopia, some farmers may use improved seed without knowing it, leading to inferior yields than conventional seed due to a lack of the application of complementary inputs.

Misclassification of improved seed has been studied in the case of cassava in Nigeria \citep{wossen2019estimating,macours2019comment}. The authors use DNA fingerprinting data to show that there is a 22\% bias in the productivity impact of improved cassava seed. In the context of seed adoption, however, there is evidence to suggest that the productivity impacts of seed affects households heterogeneously \citep{Suri2011-oi}, and that knowing average impacts to adoption may not be as informative. The observed effect of adoption may be low because households that would have a comparative advantage in adoption of improved seed are constrained by a lack of access or infrastructure. Additionally, the non-classical measurement error introducted by misclassification can bias heterogeneity in ways that may attenuate or even reverse productivity impacts. This paper provides a way to account for misclassification of seed when additional, error-free data exists, such as DNA fingerprinting data. We estimate the comparative advantage of households directly using a Group Random Coefficients (GRC) Model and correct its heterogeneity estimates using misclassification rates from DNA fingerprinting data of the same plots. \citep{Suri2011-oi, Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc}.

@@

Finally, to integrate our results together, we use the misclassification rates from the fourth round of the data and re-estimate the GRC results to be robust to misclassification \cite{michuda2021three}. We find that once we account for misclassification, heterogeneity in adoption becomes similar to other contexts: higher comparative advantage is associated with lower returns, pointing to significant barriers to adoption for those that would benefit the most from it.

@@

The article continues as follows: in Section \ref{sec:data}, we describe the data used, the sample frame, general farmer and farm characteristics, and improved seed adoption levels. Section \ref{sec:strategy} presents the methodological details of our two estimation strategies. Section \ref{sec:results} shows the estimations results, and Section \ref{sec:conclusion} presents the conclusion and implications of our analysis.

@@

\begin{figure}[H]
    \centering
    \caption{Spatial Distribution of Households Surveyed}\label{map:regions}
    \includegraphics[width=.7\textwidth]{results/figures/map_hhids.pdf}
    \vspace*{-1em}
    \begin{table}[H]
        \centering
        \begin{tabular}{p{0.6\textwidth}} 
            \begin{tablenotes}
                  \small
                  \item Note: Red markers denote the locations of households
            \end{tablenotes}
        \end{tabular}
    \end{table}  
\end{figure}

@@

\subsection{Improved seeds}

@@

Table \ref{tbl:Trajectories} maps out the frequencies of the trajectories present in the data, sorted by frequency. Most households do not adopt at all (``never-adopters''), but the second most frequent trajectory are those that adopt in every wave (``always-adopters''). Let $\mathcal{H}$ be the set of these trajectories in the data and let $\mathcal{H}_S$ be the set of switcher trajectories, i.e. $\mathcal{H}_S = \mathcal{H}\backslash\{(0,0,0), (1,1,1)\}$

@@

Once we generate $\theta$ and $\Delta$ for each trajectory, we can examine heterogeneous effects to adoption. Conditional on the LCA being satisfied, that means that there will be eight comparative advantages available for comparison, which is rich in its information, but difficult to interpret. As a result, we will aggregate $\theta$ in two ways to improve interpretability: (1) by which wave the household adopted, and (2) the number of times adoption took place. We do this by taking a simple average, grouped by each of these factors. The first aggregation will tell us whether there are any wave-specific differences in comparative advantage. The second aggregation will tell us whether adopting more times corresponds to higher comparative advantage and associated returns. In both cases, these aggregations reflect the economic landscape that drive returns to adoption.

To compare the results when using a dichotomous adoption decision with an intensity of adoption variable, we run a similar regression to the above to calculate the comparative advantage of adopting a certain number of times. We estimate an equation similar to Eq. \ref{eq:GRC_Suri}, but use the number of times that a household adopted an improved maize variety (here, $NAdopt$), not its trajectory.

\begin{align}
y_{it}&=\sum_{\underline{na}\in\mathcal{NA}\backslash 3}\nu_{\underline{h}}+\Delta_{(0,0,1)}h_{it}+\phi(\nu_{(0,0,1)}-\nu_{(0,0,1)})_{it}1\{NAdopt_{i}=1\}\nonumber\\
&~~~+\left(\nu_{(1,1,1)}+\phi\left(\nu_{(1,1,1)}-\nu_{(0,0,1)}\right)\right)h_{it}1\{NAdopt_{i}=3\} + X_{it}\beta +\varepsilon_{it},\label{eq:GRC_NAdopt}
\end{align}

Here instead of a set of trajectories, we have a set of the number of times adopted $\mathcal{NA}$. $\nu$ denotes the average yield from adoption $na$ times. We use adopting once as the base.

@@

The ATT is equal to the difference between Equation \ref{eq:att_1} and Equation \ref{eq:att_0}. The ATU is the difference between Equation \ref{eq:atu_0} and Equation \ref{eq:atu_1}. Note that the observed expectations are those linked to Equations \ref{eq:att_1} and \ref{eq:atu_0} for adopters and non-adopters, while the counterfactual expectations are those presented in Equations \ref{eq:att_0} and \ref{eq:atu_1} respectively. In the next section we report these results with bootstrapped standard errors in order to measure the yield benefit of adoption for adopters and non-adopters.

@@

Figure \ref{fig:theta_delta_raw} shows a plot of the comparative advantages and returns for the cropcuts model with controls. These results highlight two interesting puzzles: it would seem that for many trajectories, comparative advantage is either very close to 0, implying that there is no difference between farmers that adopt and not adopt. For trajectory $(0,1,1)$, comparative advantage is exceedingly negative. This is peculiar, as this would imply that farmers that adopted in both waves 2 and 3 would actually be better off not adopting, even though their returns are positive (as shown in the bottom figure). Another peculiarity in this figure is that it would seem that if we compare trajectories that only differ by one wave's adoption, comparative advantage and returns seem to fall. For example, the transition from $(0,1,0)$ to $(0,1,1)$ leads to a large fall in comparative advantage. The same can be seen less starkly for the move from $(1,0,0)$ to $(1,0,1)$ and $(1,1,0)$ to $(1,1,1)$. But despite this loss in comparative advantage, returns seem to increase. This seems to imply that adopting more often leads to higher comparative advantage, but slightly higher returns.

\begin{figure}
    \centering
    \caption{Comparative Advantage and Returns by Trajectory (Dry Cropcuts)}
    \includegraphics[scale=0.7]{results/figures/theta.png}\label{fig:theta_delta_raw}
        \vspace*{-2em}
    \begin{table}[H]
        \centering
        \begin{tabular}{p{0.8\textwidth}} 
            \begin{tablenotes}
                  \small
                  \item Note: Author's calculations using ESS. The figures shows the comparative advantage and the returns to adoption for each adoption trajectory. Dry cropcuts only
            \end{tablenotes}
        \end{tabular}
    \end{table}
\end{figure}

@@

To better understand the above puzzling results, we explore similar yield estimations using the different definitions of improved seeds from DNA fingerprinting. The DNA fingerprinting data allows us to be confident in whether a seed that was used, was in fact improved. We run similar switching regressions using different thresholds of DNA purity (70, 90 and 95\% purity), hybrid vs. open-pollinated varieties, and drought-tolerant maize (DTMZ). Table \ref{tab:switch2} shows the results.

@@

In this section, we will introduce a misclassification-robust approach to the GRC method that will allow us to take misclassification into account when calculating comparative advantage. The DNA identification from ESS4 allows us to verify how often an improved seed is actually used when a farmer self-reports adopting improved seed. The methodology in this section builds on work that was done in \cite{michuda2021three}, which used confusion matrices from machine learning algorithms as a way to estimate a causal effect that was robust to misclassification. Misclassification introduces non-classical measurement error that may attenuate estimates and bias them towards 0 in the binary case. In the multi-regime case, misclassification will tend to make estimates too close to one another, leading to smaller or even wrong-signed estimates. We construct a confusion matrix based on the DNA fingerprinting data in the fourth wave and apply it to the previous three waves.

@@

Table \ref{tab:misclassification} shows a table of different hybrid seed definitions and the percentage of times that a farmer self-reported traditional or hybrid seed, and whether they in fact were using traditional or hybrid seed, based on the DNA fingerprinting data. Table \ref{tab:a} shows that around 86\% of farmers actually use traditional seed when they self-report using traditional seed, and around 70\% of the time, farmers use hybrid seed of at least 95\% purity when they self-report using hybrid. There are in fact instances where farmers, however, misclassify their self-reports. For instance, in Table \ref{tab:b}, a farmer that self-reports using drought-tolerant maize, only in fact uses drought-tolerant maize 13\% of the time. There are similarly large misclassifications for seed released after 2010. The previous section highlighted that different definition of maize could actually lead to vastly different treatment impacts from seed adoption.

In order to apply these misclassification probabilities to the previous three waves, we made assumptions about the nature of misclassification: that the rate of misclassification is constant across time and exhibits no dynamic behavior. Although these are strong assumptions, they come from the limitations of the data; the fourth wave of the ESS was the first to include DNA fingerprinting, and so was the first where a table like the one in Table \ref{tab:misclassification} could be calculated. Let $p_{ij}$ be the $i,j$th cell of a misclassification table as in Table \ref{tab:misclassification}. Suppose that we construct the probability of being in a true trajectory $h = (i,j,k)$, intersected with being in a reported trajectory $\hat{h} = (\hat{i},\hat{j}, \hat{k})$ as:

We can use Eq. \ref{eq:misclass} to construct the joint distribution, $f(h, \hat{h})$ for each trajectory in the sample. Since Table \ref{tab:a} provides a general definition of an improved seed, we use it for the main body of the paper. Table \ref{tab:cm} shows the results of the construction of the joint distribution. The first thing to notice is that the diagonal elements tend get lower with more adoption of seed. This is due to the fact that a farmer only has a 70\% probability of having a true-positive outcome when adopting hybrid seed. With three years of adoption, the probability of having at least one year of adoption being misclassified is high. Another thing to notice is that if, for example, a farmer adopts hybrid seed for 2 out of 3 years, their misclassification probability does not change if they reported adopting in wave 1 and wave 2 or wave 2 and wave 3. Both these features of the confusion matrix have to do with the fact that these probabilities are independent of any dynamic transition or auto-correlation. It would be expected that if there was some learning or signal from buying seed from a particular seller, the misclassification probability would get smaller with more adoption. Thus, our case can be thought of as a conservative lower-bound on the effect of misclassification on comparative advantage and returns.

@@

Once we estimate the parameters, we once again calculate comparative advantage and returns to adoption. Figure \ref{fig:mr-grc} shows the results of the MR-GRC and previous GRC results side-by-side. What's striking is that once misclassification is accounted for, comparative advantage to adoption is no longer negative to trajectories that adopt hybrid. In fact, the number of adoptions is positively correlated with comparative advantage. This suggests that households are sorting based on how well they can use hybrid. What is interesting, however, is that once we account for misclassification, returns to adoption become almost universally negative. This is indicative of a negative $phi$ term or that, there is negative selection into hybrid. Similar to \cite{Suri2011-oi}, a negative $\phi$ term suggests that there may be institutional factors at play that make the returns to adoption negative and this may be the reason why households do not adopt. The only trajectory that has positive returns to adoptions is $(0,0,1)$. What's peculiar about that trajectory is that comparative advantage is closest to 0 than any other trajectory. This may suggest that all other things being equal, adoption still presents positive returns over traditional seed, especially if that adoption occurs in the third wave.

\begin{figure}
    \centering
    \includegraphics[scale=0.7]{results/figures/mr-theta.png}
    \caption{Comparative Advantage and Returns to Adoption: GRC and MR-GRC Results}
    \label{fig:mr-grc}
            \vspace*{-2em}
    \begin{table}[H]
        \centering
        \begin{tabular}{p{0.8\textwidth}} 
            \begin{tablenotes}
                  \small
                  \item Note: Figure shows GRC and MR-GRC results side-by-side. GRC results are identical as the ones from Figure \ref{fig:theta_delta_raw}. MR-GRC estimated using likelihood function in Eq. \ref{eq:mr-grc}. Misclassification defined using DNA purity data from the fourth wave of ESS. 
            \end{tablenotes}
        \end{tabular}
    \end{table}
\end{figure}

\section{Discussion}

@@

\appendix

>>>>

% 	\affil[1]{\small \emph{The World Bank}}
% 	\affil[2]{\small \emph{FAO}}
% 	\affil[3]{\small \emph{Assistant Research Professor at the Center for Data Science for Enterprise and Society, Cornell University}}
\author{%
 Aleksandr Michuda \footnote{Center for Data Science for Enterprise and Society, Cornell University}%
 \and Cristina Chiarella\footnote{Earth and Life Institute, UCLouvain, 1348 Louvain-la-Neuve, Belgium}%
 \and Oscar Barriga-Cabanillas\footnote{The World Bank}%
  \and Juan Sebastian Correa \footnote{Food and Agriculture Organization of the United Nations (FAO), 00153, Rome, Italy}%
  }

%   Assistant Research Professor at the 
	\date{\today}

    \clearpage\maketitle
	\thispagestyle{empty}
	\vspace*{-2em}
    \begin{center}\begin{abstract}
    \singlespacing
			\noindent  
			Despite large investments in  developing new  maize varieties, evidence shows that few farmers consistently adopt this new technology over time and that average effects mask considerable hetergeneity in impacts. Misclassification of seed is a crucial problem in these contexts, as non-classical measurement error can bias the impacts. We take a step further and study misclassification in the context of heterogeneous impacts to adoption. We use a three-year panel to estimate heterogeneous returns to adoption from self-reports of hybrid seed use and find peculiar results driven by bias due to the misclassification of seed. We exploit unique information on maize DNA-fingerprinting collected over the same areas in 2018. We find that misclassification in self-reported hybrid seed adoption mask heterogeneity in the actual quality of the genetic material of the seeds farmers categorized as improved varieties in the self-reported data. Positive returns to adoption are found for those farmers that use higher-purity germplasm, drought-tolerant maize, and newly released varieties. When we account for misclassification in our heterogeneity analysis, we find that the puzzle is resolved. Our findings point to the wide dispersion of older and genetically diluted varieties, for which poorer farmers may be paying a premium. The implications of our findings speak to the need for policies to better target context and geography, expand accessibility of improved seeds, and make higher yields varieties more inclusive.

@@

%%% Why it is important
% Food security is a major challenge in many East African countries, and Ethiopia is no exception, having historically struggled to provide an adequate and reliable food supply \citep{Ramakrishna2002-hv, Jaleta2018-oj}. Since the drought of 1984, Ethiopia has become increasingly reliant on maize, and as of today, maize constitutes the most widespread crop in the country, with total area cultivated increasing and average yields almost doubling since 1990.\footnote{FAO statistics show the total area cultivated with Maize in 2019 is 25\% larger than the next most widespread crop (sorghum), and 27\% and 139\% larger than the next two following crops, wheat and barley, respectively.}\footnote{Improved security conditions and maize productivity contributed with an important decrease in undernourishment which fell from 47\% in 1990 to 14\% in 2018 (The World Bank, 2021). But despite these gains, food insecurity still affects more than half the population of Ethiopia. The country has also recently seen increases in food insecurity because of COVID-19, the conflict in Ethiopia’s Tigray region and the desert locust outbreak.} But despite this progress a large productivity gap persists between the African continent and the rest of the developing world.\footnote{From 2016 to the present, cereal crop yields in Sub-Saharan Africa have consistently been around 70-80\% lower than Europe or North America \citep{worldbankyield}.} The persistence of this gap is often attributed to different adoption rates of green revolution technologies, which include improved seeds. Evidence has shown that the adoption of improved seed leads to increases in agricultural yields \citep{Carter2014-fm}, food security \citep{Shiferaw2014-op}, and poverty reduction \citep{Minten2008-tj}.\footnote{Nevertheless, critics of green revolution technologies have also claimed that these technological packages have not been as profitable for many farmers in Africa as promised. Even the most successful of these countries, Ethiopia, is not on track to meet the goals of the Alliance for a Green Revolution for Africa (AGRA). This is due to high costs, and heavy dependence on subsidies \citep{Wise20}.} But in fact, only 13\% of maize farmers in Ethiopia have adopted improved seeds consistently between 2011 and 2015. A large percentage has adopted new varieties inconsistently, and the largest share of farmers (62\%) have not used improved seeds at all during the same period.

@@

Low adoption of hybrid seed, despite its benefits, has been a longstanding question in the study of agricultural technologies. Several answers have been proposed to explain such low adoption of improved varieties. These include lack of liquidity and other economic constraints \citep{Carter2014-fm}, growing conditions such as the type of soil, input use, and farmer characteristics \citep{Munshi2004-og}, imperfections in credit markets \citep{Croppenstedt2003-pq}, ambiguously defined property rights \citep{Place2000-el}, lack of commitment devices \citep{Duflo2009-iv}, high transportation costs that increase the cost of agricultural inputs \citep{Byerlee2013-qk}, and improved seed supply constraints \citep{Bird2020-nt}. An alternative mechanism for a lack of improvement has been the misclassification of seed and seed dilution (\citealt{wossen2022misperceiving}; \citealt{kosmowski2019varietal}). Net returns to improved seed is often lower than expected in the long term because a large percentage of farmers use diluted improved seed varieties, or ones that have been recycled over many generations (\citealt{euler2022because}; \citealt{wossen2019estimating}). Evidence suggests that yield improvement and disease resistance wanes over time \citep{Abate2015-rj}. Conversely, given the ubiquity of improved seed varieties over the course of decades of access in Ethiopia, some farmers may use improved seed without knowing it, leading to inferior yields than conventional seed due to a lack of the application of complementary inputs.

Misclassification of improved seed and its implications for measuring productivity impacts has been a relatively new branch of study, with the advent of DNA fingerprinting data. For instance, \cite{wossen2019estimating} use DNA fingerprinting data to show that there is a 22\% bias in the productivity impact of improved cassava seed in Nigeria. In the context of hybrid seed adoption, however, there is evidence to suggest that productivity impacts of seed affect households heterogeneously \citep{Suri2011-oi}, and that average impacts to adoption may not be as informative. The observed effect of adoption may be low because households that would have a comparative advantage in adoption of improved seed are constrained by a lack of access to improved seed or infrastructure. Additionally, the non-classical measurement error introduced by misclassification can bias heterogeneity in ways that may attenuate or even reverse productivity impacts. This paper provides a way to account for misclassification of seed when additional, error-free data exists, such as DNA fingerprinting data. We estimate the comparative advantage of households directly using a Group Random Coefficients (GRC) Model and correct its heterogeneity estimates using misclassification rates from DNA fingerprinting data of the same plots \citep{Suri2011-oi, Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc}.

@@

Finally, to integrate our results together, we use the misclassification rates from the fourth round of the data and re-estimate the GRC results to be robust to misclassification (\citealt{michuda2021three}). We find that once we account for misclassification, heterogeneity in adoption becomes similar to other contexts: higher comparative advantage is associated with lower returns, pointing to significant barriers to adoption for those that would benefit the most from it.

@@

The article continues as follows: in Section \ref{sec:data}, we describe the data used, the sample frame, general farmer and farm characteristics, and improved seed adoption levels. Section \ref{sec:strategy} presents the methodological details of our two estimation strategies. Section \ref{sec:results} shows the estimations results, Section \ref{sec:disc} discusses our findings in light of prior evidence, and Section \ref{sec:conclusion} presents the conclusion and implications of our analysis.

@@

\begin{figure}[H]
    \centering
    \caption{Spatial Distribution of Households Surveyed}\label{map:regions}
    \includegraphics[width=.7\textwidth]{results/figures/map_hhids.pdf}
    \vspace*{-1em}
    \begin{table}[H]
        \centering
        \begin{tabular}{p{0.6\textwidth}} 
            \begin{tablenotes}
                  \small
                  \item Note: Red markers denote the locations of households. Coordinates are slightly modified and repeated for confidential reasons.
            \end{tablenotes}
        \end{tabular}
    \end{table}  
\end{figure}

@@

\subsection{Improved Seeds}

@@

Table \ref{tbl:trajectories} maps out the frequencies of the trajectories present in the data, sorted by frequency. Most households do not adopt at all (``never-adopters''), but the second most frequent trajectory are those that adopt in every wave (``always-adopters''). Let $\mathcal{H}$ be the set of these trajectories in the data and let $\mathcal{H}_S$ be the set of switcher trajectories, i.e. $\mathcal{H}_S = \mathcal{H}\backslash\{(0,0,0), (1,1,1)\}$

@@

% Once we generate $\theta$ and $\Delta$ for each trajectory, we can examine heterogeneous effects to adoption. Conditional on the LCA being satisfied, that means that there will be eight comparative advantages available for comparison, which is rich in its information, but difficult to interpret. As a result, we will aggregate $\theta$ in two ways to improve interpretability: (1) by which wave the household adopted, and (2) the number of times adoption took place. We do this by taking a simple average, grouped by each of these factors. The first aggregation will tell us whether there are any wave-specific differences in comparative advantage. The second aggregation will tell us whether adopting more times corresponds to higher comparative advantage and associated returns. In both cases, these aggregations reflect the economic landscape that drive returns to adoption.

% To compare the results when using a dichotomous adoption decision with an intensity of adoption variable, we run a similar regression to the above to calculate the comparative advantage of adopting a certain number of times. We estimate an equation similar to Eq. \ref{eq:GRC_Suri}, but use the number of times that a household adopted an improved maize variety (here, $NAdopt$), not its trajectory.

% \begin{align}
% y_{it}&=\sum_{\underline{na}\in\mathcal{NA}\backslash 3}\nu_{\underline{h}}+\Delta_{(0,0,1)}h_{it}+\phi(\nu_{(0,0,1)}-\nu_{(0,0,1)})_{it}1\{NAdopt_{i}=1\}\nonumber\\
% &~~~+\left(\nu_{(1,1,1)}+\phi\left(\nu_{(1,1,1)}-\nu_{(0,0,1)}\right)\right)h_{it}1\{NAdopt_{i}=3\} + X_{it}\beta +\varepsilon_{it},\label{eq:GRC_NAdopt}
% \end{align}

% Here instead of a set of trajectories, we have a set of the number of times adopted $\mathcal{NA}$. $\nu$ denotes the average yield from adoption $na$ times. We use adopting once as the base.

@@

The ATT is equal to the difference between Equation \ref{eq:att_1} and Equation \ref{eq:att_0}. The ATU is the difference between Equation \ref{eq:atu_0} and Equation \ref{eq:atu_1}. Note that the observed expectations are those linked to Equations \ref{eq:att_1} and \ref{eq:atu_0} for adopters and non-adopters, while the counterfactual expectations are those presented in Equations \ref{eq:att_0} and \ref{eq:atu_1} respectively. We report these results with bootstrapped standard errors in order to measure the yield benefit of adoption for adopters and non-adopters.

As a last step, we introduce a misclassification-robust approach to the GRC method, which allows us to take misclassification into account when calculating comparative advantage. Using the DNA identification from ESS4 we are able to verify how often an improved seed is actually used when a farmer self-reports adopting improved seed. The methodology in this section builds on work that was done in \cite{michuda2021three}, which used confusion matrices from machine learning algorithms as a way to estimate a causal effect that was robust to misclassification. Misclassification introduces non-classical measurement error that may attenuate estimates and bias them towards 0 in the binary case. In the multi-regime case, misclassification will tend to make estimates too close to one another, leading to smaller or even wrong-signed estimates.

Let $p_{ij}$ be the $i,j$th cell of a two by two misclassification matrix. Suppose that we construct the probability of being in a true trajectory $h = (i,j,k)$, intersected with being in a reported trajectory $\hat{h} = (\hat{i},\hat{j}, \hat{k})$ as:

We can use Eq. \ref{eq:misclass} to construct the joint distribution, $f(h, \hat{h})$ for each trajectory in the sample.

@@

Figure \ref{fig:theta_delta_raw} shows a plot of the comparative advantages and returns for the cropcuts model with controls (full results for both self-report and cropcuts are shown in the Appendix). These results highlight two interesting puzzles: it would seem that for many trajectories, comparative advantage is either very close to 0, implying that there is no difference between farmers that adopt and not adopt. For trajectory $(0,1,1)$, comparative advantage is exceedingly negative. This is peculiar, as this would imply that farmers that adopted in both waves 2 and 3 would actually be better off not adopting, even though their returns are positive (as shown in the bottom figure). Another peculiarity in this figure is that it would seem that if we compare trajectories that only differ by one wave's adoption, comparative advantage and returns seem to fall. For example, the transition from $(0,1,0)$ to $(0,1,1)$ leads to a large fall in comparative advantage. The same can be seen less starkly for the move from $(1,0,0)$ to $(1,0,1)$ and $(1,1,0)$ to $(1,1,1)$. But despite this loss in comparative advantage, returns seem to increase. This seems to imply that adopting more often leads to higher comparative advantage, but slightly higher returns.

\begin{figure}
    \centering
    \caption{Comparative Advantage and Returns by Trajectory (Dry Cropcuts)}
    \includegraphics[scale=0.7]{results/figures/theta.png}\label{fig:theta_delta_raw}
        \vspace*{-2em}
    \begin{table}[H]
        \centering
        \begin{tabular}{p{0.8\textwidth}} 
            \begin{tablenotes}
                  \small
                  \item Note: Author's calculations using ESS. The figures shows the comparative advantage and the returns to adoption for each adoption trajectory. Log(Dry cropcuts) used as outcome. Top figure denotes comparative advantage for each trajectory. Bottom figure denotes returns to adoption for each trajectory.
            \end{tablenotes}
        \end{tabular}
    \end{table}
\end{figure}

@@

To better understand the above puzzling results, we explore similar yield estimations using the different definitions of improved seed from DNA fingerprinting. The DNA fingerprinting data allows us to be confident in whether a seed that was used, was in fact improved. We run similar switching regressions using different thresholds of DNA purity (70, 90 and 95\% purity), hybrid vs. open-pollinated varieties, and drought-tolerant maize (DTMZ). Table \ref{tab:switch2} shows the results.

@@

We construct a confusion matrix based on the DNA fingerprinting data in the fourth wave and apply it to the previous three waves. Table \ref{tab:misclassification} shows a table of different hybrid seed definitions and the percentage of times that a farmer self-reported traditional or hybrid seed, and whether they in fact were using traditional or hybrid seed, based on the DNA fingerprinting data. Table \ref{tab:a} shows that around 86\% of farmers actually use traditional seed when they self-report using traditional seed, and around 70\% of the time, farmers use hybrid seed of at least 95\% purity when they self-report using hybrid. There are in fact instances where farmers, however, misclassify their self-reports. For instance, in Table \ref{tab:b}, a farmer that self-reports using drought-tolerant maize, only in fact uses drought-tolerant maize 13\% of the time. There are similarly large misclassifications for seed released after 2010. The previous section highlighted that different definition of maize could actually lead to vastly different treatment impacts from seed adoption.

@@

In order to apply these misclassification probabilities to the previous three waves, we made assumptions about the nature of misclassification: that the rate of misclassification is constant across time and exhibits no dynamic behavior. Although these are strong assumptions, they come from the limitations of the data; the fourth wave of the ESS was the first to include DNA fingerprinting, and so was the first where a table like the one in Table \ref{tab:misclassification} could be calculated. Using these missclassification probabilities, we calculate the joint distribution for each trajectory in the sample, using a purity threshold of 95\% or higher as the chosen ``true'' definition of improved seed. Table \ref{tab:cm} shows the results of the construction of the joint distribution. The first thing to notice is that the diagonal elements tend get lower with more adoption of seed. This is due to the fact that a farmer only has a 70\% probability of having a true-positive outcome when adopting hybrid seed. With three years of adoption, the probability of having at least one year of adoption being misclassified is high. Another thing to notice is that if, for example, a farmer adopts hybrid seed for 2 out of 3 years, their misclassification probability does not change if they reported adopting in wave 1 and wave 2 or wave 2 and wave 3. Both these features of the confusion matrix have to do with the fact that these probabilities are independent of any dynamic transition or auto-correlation. It would be expected that if there was some learning or signal from buying seed from a particular seller, the misclassification probability would get smaller with more adoption. Thus, our case can be thought of as a conservative lower-bound on the effect of misclassification on comparative advantage and returns.

@@

Once we estimate the parameters, we once again calculate comparative advantage and returns to adoption. Figure \ref{fig:mr-grc} shows the results of the MR-GRC and previous GRC results side-by-side. Misclassification of seed biases the estimates of the heterogeneous effects, both in terms of the estimation of comparative advantage and the returns to adoption. For instance, in the case of $(0,1,1)$, misclassification actually reverses the returns to adoption from being positive to negative. In other instances, misclassification attenuates the returns to adoption. Once misclassification is accounted for, the comparative advantage to adoption is no longer negative to trajectories that adopt hybrid seed.  But despite considerable heterogeneity in comparative advantage, returns to adoption to most trajectories is negative, but similar in magnitude. This would translate to a $\phi$ term that is close to 0, but slightly negative, meaning that the there is either no sorting of households based on comparative advantage or that there is negative sorting. This is similar to the Kenyan context in \cite{Suri2011-oi}, who finds a negative $\phi$ term and negative returns to adoption, meaning  that there are significant barriers to using hybrid seed, and that farmers who are the most poised to gain from adoption are constrained by a lack of access to seed and other complementary inputs.

\begin{figure}
    \centering
    \includegraphics[scale=0.7]{results/figures/mr-theta.png}
    \caption{Comparative Advantage and Returns to Adoption: GRC and MR-GRC Results}
    \label{fig:mr-grc}
            \vspace*{-2em}
    \begin{table}[H]
        \centering
        \begin{tabular}{p{0.8\textwidth}} 
            \begin{tablenotes}
                  \small
                  \item Note: Figure shows GRC and MR-GRC results side-by-side. GRC results are identical as the ones from Figure \ref{fig:theta_delta_raw}. MR-GRC estimated using likelihood function in Eq. \ref{eq:mr-grc}. Misclassification defined using DNA purity data from the fourth wave of ESS. Log(Dry cropcuts) used as outcome. Top figure denotes comparative advantage for each trajectory. Bottom figure denotes returns to adoption for each trajectory.
            \end{tablenotes}
        \end{tabular}
    \end{table}
\end{figure}

\section{Discussion}\label{sec:disc}

@@

\section{Appendix}
